---
title: "AD Inoculum Test Codes"
author: "Hezhou (Jenny) Ding"
date: '2022-05-17'
output: html_document
---
This file demonstrates the generations of the graphs in the manuscript. This file uses bacterial data as an example for how we generated the relative abundance graph and beta diversity graph, as well as how we ran PERMANOVA on the beta diversity. The file also shows how we calculated alpha diversity using the prokyryotic genus data. The generations of graphs for archaea genus are not shown here becasue it essentially shares the same process as bacterial data.

## Load Libraries
```{r, message=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
library(plyr)
```


# Bacteria
```{r,echo=FALSE}
setwd("C:\\Users\\Jenny Ding\\Documents\\Research\\Data\\2020 Feb AD Prelim Project\\DNA Results\\Downstream DNA results\\1Publish to Github")
```

```{r}
bac.genus.data <- read.csv("bacteria genus.csv", check.names = FALSE)
```
## Relative Abundance
### Slice the top 10-15 genus by each treatment group's
```{r}
t1 <- bac.genus.data %>% slice_max(`0-F-T-A-1`, n=10)
t2 <- bac.genus.data %>% slice_max(`0-F-T-A-2`, n=10)
t3 <- bac.genus.data %>% slice_max(`0-F-T-1`, n=10)
t4 <- bac.genus.data %>% slice_max(`0-F-T-2`, n=10)
t5 <- bac.genus.data %>% slice_max(`0-F-A-1`, n=10)
t6 <- bac.genus.data %>% slice_max(`0-F-A-2`, n=10)
t7 <- bac.genus.data %>% slice_max(`24-F-T-A-1`, n=10)
t8  <-bac.genus.data %>% slice_max(`24-F-T-A-2`, n=10)
t9 <- bac.genus.data %>% slice_max(`24-F-T-1`, n=10)
t10 <- bac.genus.data %>% slice_max(`24-F-T-2`, n=10)
t11 <- bac.genus.data %>% slice_max(`24-F-A-1`, n=10)
t12 <- bac.genus.data %>% slice_max(`24-F-A-2`, n=10)
```

### Subset the targeted treatment group from each sorted sheet
```{r}
t1a <- select(t1,`Genus`,`0-F-T-A-1`)
t2a <- select(t2, `Genus`, `0-F-T-A-2`)
t3a <- select(t3, `Genus`, `0-F-T-1`)
t4a <- select(t4, `Genus`, `0-F-T-2`)
t5a <- select(t5, `Genus`, `0-F-A-1`)
t6a <- select(t6, `Genus`, `0-F-A-2`)
t7a <- select(t7, `Genus`, `24-F-T-A-1`)
t8a <- select(t8, `Genus`, `24-F-T-A-2`)
t9a <- select(t9, `Genus`, `24-F-T-1`)
t10a <- select(t10, `Genus`, `24-F-T-2`)
t11a <- select(t11, `Genus`, `24-F-A-1`)
t12a <- select(t12, `Genus`, `24-F-A-2`)
```

### Combine the selected column using full join function in plyr
```{r}
bac.genus.top10 <- plyr::join_all(list(t1a,t2a,t3a,t4a,t5a,t6a,t7a,t8a,t9a,t10a,t11a,t12a),
                           by = "Genus", type = "full")
# Remove NA
bac.genus.top10[is.na(bac.genus.top10)] <- 0

# Generate "Others"
colSums(bac.genus.top10[,2:13]) # Sum of each column
Others <- 1- colSums(bac.genus.top10[,2:13]) # Others = 1-(sum of the members)
bac.genus.top10[32,2:13] <- Others; bac.genus.top10[32,1]<- "1Others" # Give it a name
```
### Plot the relative abundance
```{r}
# Reshape the data frame to long form so it can be used for ggplot
library(reshape2)
genus_melted <- melt(bac.genus.top10,id=("Genus"))
theme_set(theme_bw())
# Add a column for time of the samples
cbind(genus_melted,Time=c(0))
# Give values
genus_melted[193:384,"Time"]="Day24"; genus_melted[1:192,"Time"]="Day0";
# Map values into axis labels using mapvalues from plyr package and then group on that 
genus_melted$variable <- mapvalues(genus_melted$variable, 
                                   c('0-F-T-A-1','0-F-T-A-2',"0-F-T-1", "0-F-T-2","0-F-A-1", 
                                     "0-F-A-2","24-F-T-A-1","24-F-T-A-2","24-F-T-1", "24-F-T-2",
                                     "24-F-A-1", "24-F-A-2"),
                                   c("F-T-A-1","F-T-A-2","F-T-1", "F-T-2", "F-A-1", 
                                     "F-A-2","F-T-A-1","F-T-A-2","F-T-1", "F-T-2", "F-A-1", 
                                     "F-A-2"))

# Plot the stacked bar chart
## Pull the genus names
Genus.name <- sort(bac.genus.top10[,1])
Genus.name[Genus.name == "1Others"] <- "Others"

theme_set(theme_bw())
p <- ggplot(genus_melted, aes(fill=Genus, y=value, x=variable)) + 
  geom_bar(position="stack", stat="identity",color = "black", width=0.7) +
  labs(x="",y="Relative abundance") + 
  theme(axis.text.y = element_text(size=10),axis.text.x = element_text(angle=-90,size=10,margin = margin(t = 2, unit = "pt")), 
        axis.title.y = element_text(size=10, margin = margin(r = 20, unit = "pt")),
        legend.text = element_text(size=8, face = "italic"), legend.title = element_text(size=10),
        panel.border = element_rect(size = 1.2)) +
  scale_fill_manual(labels=as.factor(Genus.name),
                    values=c("#caf0f8","#0077b6","#4a4e69", "#ced4da","#bb3e03","#f4f1de","#fb8500","#ff8fa3",
                             "#ccd5ae","#606c38","#8c2f39","#419d78","#5e60ce","#dc2f02","#fcbf49","#bdb2ff",
                             "#5aa9e6","#757bc8","#fee440","#a5be00","#7b2cbf","#d88c9a","#feeafa","#588b8b",
                             "#b23a48","#3d5a80","#00b4d8", "#f2cc8f","#eb5e28","#6d597a","#b5e48c","#90a8c3",
                             "#ff477e","#ffffff"))

# Use FACET function to separate graphs
p + facet_grid(cols = vars(Time), scales="free") +
  theme(strip.text.x = element_text(size = 10),
        strip.background = element_rect(size = 1.2))

# Save graph
# ggsave("Bacteria.RA.png", plot = last_plot(), path = "", 
#        width = 13, height = 7, device='png', dpi=700)
```

## Beta Diversity
```{r, message=FALSE, results='hide'}
library(vegan)
library(ggplot2)
library(ggforce)
```

### Calculate distances and PCoA
```{r}
bac.genus.data=read.csv("bacteria genus.csv", check.names = FALSE, row.names = 1) 

# Transferring data frame to data matrix
bac.genus.matrix = data.matrix(bac.genus.data)

# Calculate Bray-Curtis dissimilarity matrix (many other beta distance models to choose)
beta.dist = vegdist(t(bac.genus.matrix), method = "bray", na.rm = TRUE)

# Calculate PCoA or so called MDS (very confusing stats analysis)
pcoa.data=cmdscale(beta.dist, eig=TRUE,x.ret=FALSE)
pcoa.var=round(pcoa.data$eig/sum(pcoa.data$eig)*100,1)

# Converting data to ggplot format
pcoa.value = pcoa.data$points
Sample <- c("F-T-A","F-T-A","F-T","F-T","F-A","F-A","F-T-A","F-T-A","F-T","F-T","F-A","F-A")
X <- as.numeric(pcoa.value[,1])
Y <- as.numeric(pcoa.value[,2])
pcoa.df1 <- data.frame(Sample, X, Y)

# Give values                      
cbind(pcoa.df1,Time=c(0))
pcoa.df1[1:6,"Time"]="Day0"; pcoa.df1[7:12,"Time"]="Day24";
pcoa.df1
```

### Plot
```{r}
theme_set(theme_bw())
ggplot(data=pcoa.df1, aes(x=X, y=Y))+
  geom_point(size=7,aes(shape=Sample, fill=Time),stroke=1.5)+
  scale_shape_manual(values = c(21, 22, 24) ) +
  scale_fill_manual(values = c("#f3722c","#90be6d"))+
  guides(fill = guide_legend(override.aes = list(shape = 21) ),
         shape = guide_legend(override.aes = list(fill = "black")))+
  #ellipse are used for visualization. could do "type" for distribution, or level, etc.
  theme(axis.text.y = element_text(size=16), axis.text.x = element_text(size=16),
        axis.title.x = element_text(size=18, margin = margin(t = 20, unit = "pt")),
        axis.title.y = element_text(size=18, margin = margin(r = 20, unit = "pt")),
        legend.text = element_text(size=12), legend.title = element_text(size=16), legend.key.size = unit(1, "cm"),
        panel.border = element_rect(size = 1.2)) +
  coord_fixed(xlim = c(-0.68,0.5),ylim = c(-0.4,0.55)) +
  xlab(paste("PCoA1 - ", pcoa.var[1], "%", sep=""))+
  ylab(paste("PCoA2 - ", pcoa.var[2], "%", sep = ""))+
  ggtitle("Bacterial Beta Diversity") + theme(plot.title = element_text(size = 20,margin = margin(b = 10, unit = "pt")))+ 
  geom_mark_ellipse(aes(fill = Time))

#Save file
# ggsave("Bacteria.beta.png", plot = last_plot(), path = "", 
#        width = 8, height = 8, device='png', dpi=700)

```
## PERMANOVA for Beta Diversity
#### Groups the treatments to Day0 and Day24-first 6 is day0, second 6 is day 24.
```{r, message=FALSE, results='hide'}
library(RRPP)
```

### We can use two methods for PERMANOVA
```{r}
# Import metadata
SAM <- read.csv("metadata.csv", row.names = 1)
```
#### Method1
```{r}
groups <- factor(c(rep("Day0", 6), rep("Day24", 6)))
genus.adonis <- adonis(beta.dist ~ groups, permutations = 999, method = "bray") 
print(genus.adonis)
```

#### Method2 
```{r}
otu.rrpp.d <- RRPP::lm.rrpp(beta.dist ~ Day,
                            data=SAM, SS.type="III", 
                            print.progress = FALSE, 
                            seed="random", 
                            iter=999)
anova(otu.rrpp.d, effect.type = "F")
```

# Prokaryotes
## Alpha Diversity
```{r, echo=FALSE}
setwd("C:\\Users\\Jenny Ding\\Documents\\Research\\Data\\2020 Feb AD Prelim Project\\DNA Results\\Downstream DNA results\\1Publish to Github")
```

```{r, message=FALSE, results='hide'}
library(phyloseq)
library(vegan)
```
### We can use two methods
#### Method 1: Use phyloseq
```{r}
pro.genus.data <- read.csv("prokaryote genus counts.csv", check.names = FALSE, row.names = 1) 
ASV <- pro.genus.data
row.names(ASV) <- paste0("ASV", 1:297)

Genus <- row.names(pro.genus.data)
taxa <- data.frame(Genus)
row.names(taxa) <- row.names(ASV)

# phyloseq only accept counts
ps <- phyloseq(otu_table(ASV, taxa_are_rows = TRUE),
               tax_table(as.matrix(taxa)))

alpha <- phyloseq::estimate_richness(ps, split=TRUE, measures=c("Observed", "Chao1", "Shannon", "Simpson")) 
row.names(alpha) <- colnames(ASV)
alpha <- alpha[-1,]
alpha
```

#### Method 2: Use Vegan
```{r}
# Vegan accepts compositional data
diversity(t(pro.genus.data), index = "shannon")
diversity(t(pro.genus.data), index = "simpson")
```

### ANOVA for alpha diversity
```{r}
alpha.ano <- alpha
Treatment <- c("F_T_A_0","F_T_A_0","F_T_0","F_T_0","F_A_0","F_A_0","F_T_A_24","F_T_A_24","F_T_24","F_T_24","F_A_24","F_A_24")
alpha.ano[,6] <- Treatment
alpha.ano.0 <- alpha.ano [1:6,]
alpha.ano.24 <- alpha.ano [7:12,]
```

```{r}
## ANOVA and psot HOC for day 0
library(agricolae)
ANOVA_CHAO=aov(Chao1~V6, data=alpha.ano.0)
summary(ANOVA_CHAO)
Tukey_CHAO=TukeyHSD(ANOVA_CHAO)
Tukey_CHAO
HSD.test(ANOVA_CHAO, "V6", group=T)$groups 

ANOVA_SHAN=aov(Shannon~V6, data=alpha.ano.0)
summary(ANOVA_SHAN)
Tukey_SHAN=TukeyHSD(ANOVA_SHAN)
Tukey_SHAN
HSD.test(ANOVA_SHAN, "V6", group=T)$groups 

ANOVA_SIM=aov(Simpson~V6, data=alpha.ano.0)
summary(ANOVA_SIM)
Tukey_SIM=TukeyHSD(ANOVA_SIM)
Tukey_SIM
HSD.test(ANOVA_SIM, "V6", group=T)$groups

## ANOVA and psot HOC for day 24
library(agricolae)
ANOVA_CHAO=aov(Chao1~V6, data=alpha.ano.24)
summary(ANOVA_CHAO)
Tukey_CHAO=TukeyHSD(ANOVA_CHAO)
Tukey_CHAO
HSD.test(ANOVA_CHAO, "V6", group=T)$groups 

ANOVA_SHAN=aov(Shannon~V6, data=alpha.ano.24)
summary(ANOVA_SHAN)
Tukey_SHAN=TukeyHSD(ANOVA_SHAN)
Tukey_SHAN
HSD.test(ANOVA_SHAN, "V6", group=T)$groups 

ANOVA_SIM=aov(Simpson~V6, data=alpha.ano.24)
summary(ANOVA_SIM)
Tukey_SIM=TukeyHSD(ANOVA_SIM)
Tukey_SIM
HSD.test(ANOVA_SIM, "V6", group=T)$groups  
```























 







